---
title: "combining_subquestions"
output: html_document
---

I want to create an index that describes the potential value of this year's first round draft pick versus next year's first round draft pick.

inputting the data...
```{r}
predicted_pick_value <- read_csv("../Data/predictingPickValue.csv")
pick_order_probabilities <- read_csv("../Data/pickOrderProbabilities.csv")
```
```{r}
library(tidyverse)
```

I need to combine those two datasets into one...
```{r}
predicted_pick_value <- predicted_pick_value %>%
  select(Overall, predictedAvgPSperSeason) %>%
  rename(possible_pick = Overall) %>%
  filter(possible_pick < 32) %>%
  distinct()
```

```{r}
next_year_pick_value <- left_join(pick_order_probabilities,
                                  predicted_pick_value, 
                                  by = "possible_pick")
```
I will create a new column that is the probability of getting that pick times the average PS per season of the pick-- like a probability value...

```{r}
next_year_pick_value <- next_year_pick_value %>%
  mutate(value_prob = probability * predictedAvgPSperSeason)
```
and the "value" of next year's first round pick is the sum of those probability-values...
```{r}
next_year_pick_value %>%
  summarize(value = sum(value_prob))
```

So next year's first round draft pick can be given a "value" of 2.278906

and I need to compute this year's value...
```{r}
7/8 * next_year_pick_value$predictedAvgPSperSeason[15] +
  1/8 * next_year_pick_value$predictedAvgPSperSeason[1]
```
so this year's first round draft pick can be given a value of 2.968304

This year's draft pick is more valuable than next year's so the Penguins should keep their 2020 first round pick and instead give their 2021 first-round pick.
