---
title: "data_wrangling"
output: html_document
---

This will be an R markdown to document what I do to the datasets in the hopes of getting them the way I want them...

__________________________________________________________________________
Question 1: what is the value of a draft pick? What is the relationship between pick order and success in the NHL?

(using draft pick data and player career data)
```{r}
library(tidyverse)
player_season_data <- read_csv("../Data/NHLPlayerSeasonData0506to1920.csv")
```
so that's the player-season data I got from hockey-reference.com, of all the players active in each season between 05-06 and 19-20

now I want to try to rearrange that data so that it has one column for each player and shows their career stats

```{r}
per_season_career_data <- as.data.frame(unique(player_season_data$Player)) %>%
  rename(player_name = 'unique(player_season_data$Player)')

new_var <- c(0)

per_season_career_data %>%
  mutate(gp = new_var)
  
player_season_data %>%
  filter(Player == per_season_career_data$player_name) %>%
  summarize(gp = mean(GP))

```

ok that is not working the way I want it to...
I think I'll try it as a loop, if I can figure out how to write a loop in R...
first I need to create a list of player names...

```{r}
list_of_player_names <- as.data.frame(unique(player_season_data$Player)) %>%
  rename(player_name = 'unique(player_season_data$Player)')
```

then I want to define the loop that says "for each of these names, find the average of the stats for the rows containing that name"

```{r}
for (i in length(list_of_player_names)) {
  name <- list_of_player_names[,i]
  
  per_season_career_data <- as.data.frame(name)
  
  condense_stats(name)
}
```

```{r}
condense_stats <- function (name, new_dataset = per_season_career_data,
                      old_dataset = player_season_data) {
  just_this_player <- old_dataset %>%
    filter(Player == name)
  
  per_season_career_data$gp <- 0
  
  new_dataset$gp[i] <- mean(just_this_player$GP)
}
```
so this is not working the way I want it to either, I think I will probably give up on this and go with the data that Jill found instead...

wait I found a function in dplyr that might work...

```{r}
per_season_career_data <- player_season_data %>%
  group_by(Player) %>%
  summarize(gp = sum(GP), goals = sum(G), assists = sum(A), 
            points = sum(PTS), plusminus = sum(PlusMinus), PIM = sum(PIM),
            pointShares = sum(PS), evenStrengthG = sum(EvenStrengthG),
            powerPlayG = sum(PowerPlayG), shortHandG = sum(ShortHandG),
            gameWinningG = sum(GameWinningG), 
            evenStrengthA = sum(EvenStrengthA), powerPlayA = sum(PowerPlayA),
            shortHandA = sum(ShortHandA), shots = sum(S),
            TOI = sum(TOI), blocks = sum(BLK), hits = sum(HIT), 
            foWins = sum(FOW), foLosses = sum(FOL), 
            CorsiFor = sum(CF), CorsiAgainst = sum(CA),
            FenwickFor = sum(FF), FenwickAgainst = sum(FA), 
            oiSHperc = mean(oiSHperc), oiSVperc = mean(oiSVperc), 
            oZSperc = mean(oZSperc), dZSperc = mean(dZSperc), 
            #TOIper60 = mean(TOIper60),
            #TOIatEvenStrength = mean(TOIatEvenStrength),
            takeaways = sum(TK), giveaways = sum(GV), 
            shotAttempts = sum(Sattempts), thruperc = mean(Thruperc))
```

it doesn't like the time variables, so I'm just leaving those out

now for the ones that required mutate rather than summarize

```{r}
per_season_career_data <- per_season_career_data %>%
  mutate(shotPerc = goals / shots, averageTOI = TOI / gp,
         foPerc = foWins / (foWins + foLosses), 
         CorsiPerc = CorsiFor / (CorsiFor + CorsiAgainst),
         FenwickPerc = FenwickFor / (FenwickFor + FenwickAgainst),
         PDO = oiSHperc + oiSVperc)

player_career_data <- per_season_career_data
```

that works and does what I want! hallelujah!

I want to save it as a .csv just so I have it, just in case...
```{r}
write_csv(player_career_data, "../Data/playerCareerData0506to1920.csv")
```
so now we need to combine with the draft data!

```{r}
draft_data <- read_csv("../Data/NHLdraftdata2005-2016.csv")
```
now we need to match players' career data to their draft data based on their name and player ID

first isolate the player ID

```{r}
draft_data <- draft_data %>%
  mutate(ID = gsub(".*\\\\", " ", Player))
player_career_data <- player_career_data %>%
  mutate(ID = gsub(".*\\\\", " ", Player))
```
now we have to make sure that we only have the columns that we actually want, and that none of the columns have the same name

```{r}
draft_data_clean <- draft_data %>%
  select(Year, Round, Overall, Team, Player, Nat., Pos, Age, To, 
         YrsPlayed, AmateurTeam, AmateurLeague, ID) %>%
  rename(draft_year = Year)

player_career_data_clean <- player_career_data %>%
  select(-Player)
```

now we can actually combine them using a left join, which will make players that were drafted but never played have NAs in the NHL stats columns, and will make those who played in the NHL but were never drafted (or were drafted pre-2005) disappear...

```{r}
full_draft_data_w_NAs <- left_join(draft_data_clean,
                                   player_career_data_clean,
                             by = "ID")
```
so that dataset has NAs for the players who were drafted but never played, and we want to replace those NAs with zeros to include them in the analysis...

```{r}
full_draft_data <- full_draft_data_w_NAs %>%
  filter(!is.na(Nat.))
full_draft_data[is.na(full_draft_data)] <- 0
```
and I want to save this as a csv too just to have it...

```{r}
write_csv(full_draft_data, "../Data/draftAndCareerData0506to1920.csv")
```

so that is the dataset we will use to answer that question...

_________________________________________________________________________
Question 2: Can we predict what our pick order spot will be next year?
  2.1: What is the relationship between season performance and draft pick order?
  2.2: How can we use team-season data in the season n-1 to predict performance in season n?
  
so to answer question 2.1, we need information on past draft orders, and information on past team-season performance

```{r}
simple_team_season_data <- 
  read_csv("../Data/teamSeasonDataSimple0809to1819.csv")
draft_order_data <- read_csv("../Data/draftOrder09to19.csv")
```
first I need to do some renaming to make Atlanta be Winnipeg the whole time because that makes sense for tracking over time since it's essentially the same team but with different names

```{r}
draft_order_data$Team <- ifelse(draft_order_data$Team == "ATL", 
                                "WPG", draft_order_data$Team)

simple_team_season_data$Team <- ifelse(simple_team_season_data$Team == "ATL",
                                  "WPG", simple_team_season_data$Team)
```


but then I will need to create a column in each that describes the team and the end of the season year/draft year

```{r}
draft_order_data <- draft_order_data %>%
  rename(tm = Team) %>%
  mutate(tm_season = str_c(tm, Year, sep = "_"))

simple_team_season_data <- simple_team_season_data %>%
  mutate(season_end = Season)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2008-09", 2009,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2009-10", 2010,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2010-11", 2011,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2011-12", 2012,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2012-13", 2013,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2013-14", 2014,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2014-15", 2015,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2015-16", 2016,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2016-17", 2017,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2017-18", 2018,
         simple_team_season_data$season_end)
simple_team_season_data$season_end <- 
  ifelse(simple_team_season_data$Season == "2018-19", 2019,
         simple_team_season_data$season_end)
simple_team_season_data <- simple_team_season_data %>%
  mutate(tm_season = str_c(Team, season_end, sep = "_"))

```
so they match, now to join them

```{r}
season_to_draft_order_data <- left_join(simple_team_season_data, 
                                        draft_order_data,
                                        by = "tm_season")
```
some of those column names need some cleaning
```{r}
season_to_draft_order_data <- season_to_draft_order_data %>%
  rename(PTSperc = `PTS%`) %>%
  rename(RPtperc = `RPt%`) %>%
  rename(RgPtperc = `RgPt%`)
```


then I can save it in a csv as well

```{r}
write_csv(season_to_draft_order_data,
          "../Data/seasonToDraftOrderData0809to1819.csv")
```

So switching gears again, I need to manipulate the dataset that Jill built with the complex team season data to predict performance in season n+1 from stats in season n

```{r}
complex_team_season_data <-
  read_csv("../Data/teamSeasonDataComplex0809to1920.csv")
```
```{r}
complex_team_season_data <- complex_team_season_data %>%
  mutate(season_end = season)

complex_team_season_data$team <-
  ifelse(complex_team_season_data$team == "ATL", "WPG",
         complex_team_season_data$team)

complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2008-09", 2009,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2009-10", 2010,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2010-11", 2011,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2011-12", 2012,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2012-13", 2013,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2013-14", 2014,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2014-15", 2015,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2015-16", 2016,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2016-17", 2017,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2017-18", 2018,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2018-19", 2019,
         complex_team_season_data$season_end)
complex_team_season_data$season_end <- 
  ifelse(complex_team_season_data$season == "2019-20", 2020,
         complex_team_season_data$season_end)

complex_team_season_data <- complex_team_season_data %>%
  mutate(tm_season = str_c(team, season_end, sep = "_"))

```

so the dataset with all of the stats is set up, now I need to make a second dataset containing next year's points percentage and match them up

```{r}
points_percent_nplus1 <- complex_team_season_data %>%
  dplyr::select(team, season_end, `P%`)
points_percent_nplus1$season_end <- 
  as.numeric(points_percent_nplus1$season_end) - 1

points_percent_nplus1 <- points_percent_nplus1 %>%
  mutate(tm_season = str_c(team, season_end, sep = "_"))


mismatched <- points_percent_nplus1 %>%
  dplyr::select(tm_season, `P%`) %>%
  rename(next_year_pts_perc = `P%`)

points_percent_nplus1 <- left_join(complex_team_season_data, mismatched,
                                   by = "tm_season")

```

and I want to save it as a csv...

```{r}
write_csv(points_percent_nplus1, "../Data/predictingPpercNextSeason.csv")
```

