---
title: "data_wrangling"
output: html_document
---

This will be an R markdown to document what I do to the dataset in the hopes of getting it the way I want it...

```{r}
library(tidyverse)
player_season_data <- read_csv("../Data/NHLPlayerSeasonData0506to1920.csv")
```
so that's the player-season data I got from hockey-reference.com, of all the players active in each season between 05-06 and 19-20

now I want to try to rearrange that data so that it has one column for each player and shows their career stats

```{r}
per_season_career_data <- as.data.frame(unique(player_season_data$Player)) %>%
  rename(player_name = 'unique(player_season_data$Player)')

new_var <- c(0)

per_season_career_data %>%
  mutate(gp = new_var)
  
player_season_data %>%
  filter(Player == per_season_career_data$player_name) %>%
  summarize(gp = mean(GP))

```

ok that is not working the way I want it to...
I think I'll try it as a loop, if I can figure out how to write a loop in R...
first I need to create a list of player names...

```{r}
list_of_player_names <- as.data.frame(unique(player_season_data$Player)) %>%
  rename(player_name = 'unique(player_season_data$Player)')
```

then I want to define the loop that says "for each of these names, find the average of the stats for the rows containing that name"

```{r}
for (i in length(list_of_player_names)) {
  name <- list_of_player_names[,i]
  
  per_season_career_data <- as.data.frame(name)
  
  condense_stats(name)
}
```

```{r}
condense_stats <- function (name, new_dataset = per_season_career_data,
                      old_dataset = player_season_data) {
  just_this_player <- old_dataset %>%
    filter(Player == name)
  
  per_season_career_data$gp <- 0
  
  new_dataset$gp[i] <- mean(just_this_player$GP)
}
```
so this is not working the way I want it to either, I think I will probably give up on this and go with the data that Jill found instead...

wait I found a function in dplyr that might work...

```{r}
per_season_career_data <- player_season_data %>%
  group_by(Player) %>%
  summarize(gp = sum(GP), goals = sum(G), assists = sum(A), 
            points = sum(PTS), plusminus = sum(PlusMinus), PIM = sum(PIM),
            pointShares = sum(PS), evenStrengthG = sum(EvenStrengthG),
            powerPlayG = sum(PowerPlayG), shortHandG = sum(ShortHandG),
            gameWinningG = sum(GameWinningG), 
            evenStrengthA = sum(EvenStrengthA), powerPlayA = sum(PowerPlayA),
            shortHandA = sum(ShortHandA), shots = sum(S),
            TOI = sum(TOI), blocks = sum(BLK), hits = sum(HIT), 
            foWins = sum(FOW), foLosses = sum(FOL), 
            CorsiFor = sum(CF), CorsiAgainst = sum(CA),
            FenwickFor = sum(FF), FenwickAgainst = sum(FA), 
            oiSHperc = mean(oiSHperc), oiSVperc = mean(oiSVperc), 
            oZSperc = mean(oZSperc), dZSperc = mean(dZSperc), 
            #TOIper60 = mean(TOIper60),
            #TOIatEvenStrength = mean(TOIatEvenStrength),
            takeaways = sum(TK), giveaways = sum(GV), 
            shotAttempts = sum(Sattempts), thruperc = mean(Thruperc))
```

it doesn't like the time variables, so I'm just leaving those out

now for the ones that required mutate rather than summarize

```{r}
per_season_career_data <- per_season_career_data %>%
  mutate(shotPerc = goals / shots, averageTOI = TOI / gp,
         foPerc = foWins / (foWins + foLosses), 
         CorsiPerc = CorsiFor / (CorsiFor + CorsiAgainst),
         FenwickPerc = FenwickFor / (FenwickFor + FenwickAgainst),
         PDO = oiSHperc + oiSVperc)
```



