---
title: "EDA_work"
output: html_document
---

this will be a document for initial EDA work for the three questions

first I can read in the datasets from the csv files I saved them in from the wrangling step

```{r}
library(tidyverse)

drafted_players_data <- read_csv("../Data/draftAndCareerData0506to1920.csv")

season_draft_data <- read_csv("../Data/seasonToDraftOrderData0809to1819.csv")

complex_team_season_data <-
  read_csv("../Data/predictingPpercNextSeason.csv")
```
also need the tidyverse, as usual
```{r}
library(tidyverse)
```

I guess I'll focus on the subquestion involving predicting draft order from season performance

```{r}
season_draft_data %>%
  ggplot(aes(x = PTSperc, y = PickOrder)) +
  geom_point() +
  labs(x = "Point Percentage", y = "First Round Pick Order", 
       title = "Season Performance Predicts First Round Pick Order... 
       Sort Of") +
  theme_bw()
```
what about season n performance predicting performance in season n+1 (just for the purposes of EDA for the preliminary presentation)

```{r}
complex_team_season_data %>%
  ggplot(aes(x = `P%`, y = next_year_pts_perc)) +
  geom_point() +
  labs(x = "Points Percentage in Season n", 
       y = "Points Percentage in Season n + 1", 
       title = "Teams are relatively consistent year-to-year") +
  theme_bw()
```
is there a better relationship there with another stat?
```{r}
complex_team_season_data %>%
  ggplot(aes(x = xGoalsPercentage, y = next_year_pts_perc)) +
  geom_point()
```
that actually does look kind of better, but it's a bit weird that the values are so restricted...

