---
title: "modeling_pick_order"
output: html_document
---

setup
```{r}
library(tidyverse)
library(caret)
library(mgcv)

season_draft_data <- read_csv("../Data/seasonToDraftOrderData0809to1819.csv")

```


I want to create a model to predict pick order based on season performance, as measured by points percent

first I need to create the folds for 10-fold cross validation
```{r}
set.seed(2020)

train_i <- createDataPartition(y = season_draft_data$PickOrder,
                               p = 0.9, list = FALSE) %>%
  as.numeric()

train_pickorder_data <- season_draft_data[train_i,]
test_pickorder_data <- season_draft_data[-train_i,]

```
then I will try to use the caret package to build and test models of different types

simple linear regression
```{r}
set.seed(2000)
init_linear_pickorder_train <-
  train(PickOrder ~ PTSperc,
        data = season_draft_data, method = "lm", #can use lots of methods!
        trControl = trainControl("cv", number = 10),
        preProcess = c("center", "scale"))

init_linear_pickorder_train$results
```


K nearest neighbors
```{r}
set.seed(2002)
tune_knn_pickorder_train <-
  train(PickOrder ~ PTSperc,
        data = season_draft_data, method = "knn", #can use lots of methods!
        trControl = trainControl("cv", number = 10),
        preProcess = c("center", "scale"),
        tuneGrid = expand.grid(k = 2:40)) #pick manually, but can't use tibble
#expand.grid is really useful and powerful, creates all combos

plot(tune_knn_pickorder_train)
```
so it looks like the best is with k = 29
```{r}
tune_knn_pickorder_train$bestTune
```

```{r}
tune_knn_pickorder_train$results[28,]
```


GAM
```{r}
set.seed(2008)
init_gam_pickorder_train <-
  train(PickOrder ~ PTSperc,
        data = season_draft_data, method = "gam", #can use lots of methods!
        trControl = trainControl("cv", number = 10),
        preProcess = c("center", "scale"),
        tuneGrid = data.frame(method = "GCV.Cp", select = FALSE))
init_gam_pickorder_train$results
```

Gradient boosted machine

```{r}
xgb_tune_grid <- 
  expand.grid(nrounds = seq(from = 20, to = 200, by = 20),
              eta = c(0.025, 0.05, 0.1, 0.3, 0.6),
              max_depth = c(1:4),
              gamma = 0, colsample_bytree = 1, #leave all vars
              min_child_weight = 1, #same as node size
              subsample = 1) #let all data be included

xgb_tune_control <- trainControl(method = "cv", number = 10,
                                 verboseIter = FALSE)

set.seed(2001)
init_xgb_tune_pickorder_train <- 
  train(x = dplyr::select(season_draft_data, PTSperc),
                  y = season_draft_data$PickOrder,
                  trControl = xgb_tune_control,
                  tuneGrid = xgb_tune_grid,
                  method = "xgbTree",
                  verbose = TRUE)
```
```{r}
init_xgb_tune_pickorder_train$bestTune
```

```{r}
init_xgb_tune_pickorder_train$results[162,]
```

so I have all the models made, but now I want to figure out what their out-of-sample RMSE is and show those as a table or chart

linear
```{r}
pickorder_linear_test_preds <- predict(init_linear_pickorder_train, test_pickorder_data)
RMSE(pickorder_linear_test_preds, test_pickorder_data$PickOrder)

```
KNN
```{r}
pickorder_knn_test_preds <- predict(tune_knn_pickorder_train, test_pickorder_data)
RMSE(pickorder_knn_test_preds, test_pickorder_data$PickOrder)
```
GAM
```{r}
pickorder_gam_test_preds <- predict(init_gam_pickorder_train, test_pickorder_data)
RMSE(pickorder_gam_test_preds, test_pickorder_data$PickOrder)
```

GBM
```{r}
pickorder_xgb_test_preds <- predict(init_xgb_tune_pickorder_train, test_pickorder_data)
RMSE(pickorder_xgb_test_preds, test_pickorder_data$PickOrder)
```
make a chart...
```{r}
pickorder_rmse <- as_tibble(NULL)

pickorder_rmse <- bind_rows( c(model = "Linear", 
                              RMSE = RMSE(pickorder_linear_test_preds,
                                          test_pickorder_data$PickOrder)),
                            c(model = "K Nearest Neighbors", 
                              RMSE = RMSE(pickorder_knn_test_preds,
                                     test_pickorder_data$PickOrder)),
                            c(model = "GAM", 
                              RMSE = RMSE(pickorder_gam_test_preds,
                                          test_pickorder_data$PickOrder)),
                            c(model = "Gradient Boosted", 
                              RMSE = RMSE(pickorder_xgb_test_preds,
                                          test_pickorder_data$PickOrder)))

pickorder_rmse_chart <- pickorder_rmse %>%
  mutate(RMSE = as.numeric(RMSE)) %>%
  ggplot(aes(x = model, y = RMSE)) +
  geom_point() + 
  coord_flip() +
  labs(y = "RMSE", x = "Model Type", 
       title = "Holdout error rate among tuned models") +
  theme_bw()
pickorder_rmse_chart
```
So the best model is the GAM!

