---
title: "predicting_next_season"
output: html_document
---

now we want to use our random forest to create a distribution of possible season outcomes

I need to basically re-create the random forest model from before, so I need to first manipulate the data like it was in the modeling script

```{r}
library(tidyverse)
library(caret)
library(mgcv)
library(ranger)

complex_team_season_data <-
  read_csv("../Data/predictingPpercNextSeason.csv")
complex_team_season_data <- complex_team_season_data %>%
  mutate(xGperc = xGoalsFor / (xGoalsFor + xGoalsAgainst))
```

Now I want to work on modeling next season's win percentage based on this season's stats...
I think I need to do some data manipulation to turn these into per-game stats and remove some of the named data
```{r}
next_year_model_data <- complex_team_season_data %>%
  select(-team, -season, -season_end, -tm_season) %>%
  mutate(xGFor_per_game = xGoalsFor / GP, 
         shotsFor_per_game = shotsOnGoalFor / GP,
         shotAttemptsFor_per_game = shotAttemptsFor / GP,
         goalsFor_per_game = goalsFor / GP,
         penaltiesFor_per_game = penaltiesFor / GP,
         faceoffsWonFor_per_game = faceOffsWonFor / GP,
         hitsFor_per_game = hitsFor / GP,
         takeawaysFor_per_game = takeawaysFor / GP,
         giveawaysFor_per_game = giveawaysFor / GP,
         xGAgainst_per_game = xGoalsAgainst / GP,
         shotsAgainst_per_game = shotsOnGoalAgainst / GP,
         shotAttemptsAgainst_per_game = shotAttemptsAgainst / GP,
         goalsAgainst_per_game = goalsAgainst / GP,
         penaltiesAgainst_per_game = penaltiesAgainst / GP,
         faceoffsWonAgainst_per_game = faceOffsWonAgainst / GP,
         hitsAgainst_per_game = hitsAgainst / GP,
         takeawaysAgainst_per_game = takeawaysAgainst / GP,
         giveawaysAgainst_per_game = giveawaysAgainst / GP) %>%
  select(-GP, -W, -L, -OT, -P, -RW, -ROW, -`S/O Win`, -xGoalsPercentage,
         -xGoalsFor, -shotsOnGoalFor, -shotAttemptsFor, -goalsFor, 
         -penaltiesFor, -faceOffsWonFor, -hitsFor, -takeawaysFor, 
         -giveawaysFor, -xGoalsAgainst, -shotsOnGoalAgainst, 
         -shotAttemptsAgainst, -goalsAgainst, -penaltiesAgainst, 
         -faceOffsWonAgainst, -hitsAgainst, -takeawaysAgainst, 
         -giveawaysAgainst, -shotsFor_per_game, -goalsFor_per_game,
         -shotsAgainst_per_game, -goalsAgainst_per_game,
         -penaltiesFor_per_game, -penaltiesAgainst_per_game) %>%
  rename(PTSperc = `P%`, goalsFor_per_game = `GF/GP`,
         goalsAgainst_per_game = `GA/GP`, PP_perc = `PP%`,
         PK_perc = `PK%`, netPP_perc = `Net PP%`, netPK_perc = `Net PK%`,
         shotsFor_per_game = `Shots/GP`, shotsAgainst_per_game = `SA/GP`) %>%
  filter(!is.na(next_year_pts_perc))
  
#adding thru%for and against
next_year_model_data <- next_year_model_data %>%
  mutate(shot_thru_perc_for = shotsFor_per_game / shotAttemptsFor_per_game,
         shot_thru_perc_against = 
           shotsAgainst_per_game / shotAttemptsAgainst_per_game)
```



I want to run a number of models, tuning each as needed

first I need to create the folds for 10-fold cross validation
```{r}
set.seed(2020)

train_i <- createDataPartition(y = next_year_model_data$next_year_pts_perc,
                               p = 0.9, list = FALSE) %>%
  as.numeric()

train_nextyear_data <- next_year_model_data[train_i,]
test_nextyear_data <- next_year_model_data[-train_i,]

```

then build the model...
Random Forest
```{r}
rf_tune_grid <-
  expand.grid(mtry = seq(3, 18, by = 3),
              splitrule = "variance",
              min.node.size = 5)

set.seed(2010)
init_rf_nextyear_train <-
  train(next_year_pts_perc ~ ., data = train_nextyear_data,
        method = "ranger", num.trees = 500, 
        trControl = trainControl(method = "cv", number = 5),
        tuneGrid = rf_tune_grid)
plot(init_rf_nextyear_train)
```

```{r}
init_rf_nextyear_train$bestTune
```

```{r}
init_rf_nextyear_train$results[3,]
```


so now I want to build a model using the ranger package and these tuned parameters
```{r}
final_rf_next_year <- ranger(next_year_pts_perc ~ .,
                             data = next_year_model_data,
                             splitrule = "variance",
                             min.node.size = 5,
                             mtry = 9,
                             num.trees = 500)
```

so I have the model, and now I need to get predictions from this model for the 2020-21 season

first I want a dataset that is just the current season...
```{r}
just_this_season <- complex_team_season_data

just_this_season <- just_this_season %>%
  mutate(xGFor_per_game = xGoalsFor / GP, 
         shotsFor_per_game = shotsOnGoalFor / GP,
         shotAttemptsFor_per_game = shotAttemptsFor / GP,
         goalsFor_per_game = goalsFor / GP,
         penaltiesFor_per_game = penaltiesFor / GP,
         faceoffsWonFor_per_game = faceOffsWonFor / GP,
         hitsFor_per_game = hitsFor / GP,
         takeawaysFor_per_game = takeawaysFor / GP,
         giveawaysFor_per_game = giveawaysFor / GP,
         xGAgainst_per_game = xGoalsAgainst / GP,
         shotsAgainst_per_game = shotsOnGoalAgainst / GP,
         shotAttemptsAgainst_per_game = shotAttemptsAgainst / GP,
         goalsAgainst_per_game = goalsAgainst / GP,
         penaltiesAgainst_per_game = penaltiesAgainst / GP,
         faceoffsWonAgainst_per_game = faceOffsWonAgainst / GP,
         hitsAgainst_per_game = hitsAgainst / GP,
         takeawaysAgainst_per_game = takeawaysAgainst / GP,
         giveawaysAgainst_per_game = giveawaysAgainst / GP) %>%
  select(-GP, -W, -L, -OT, -P, -RW, -ROW, -`S/O Win`, -xGoalsPercentage,
         -xGoalsFor, -shotsOnGoalFor, -shotAttemptsFor, -goalsFor, 
         -penaltiesFor, -faceOffsWonFor, -hitsFor, -takeawaysFor, 
         -giveawaysFor, -xGoalsAgainst, -shotsOnGoalAgainst, 
         -shotAttemptsAgainst, -goalsAgainst, -penaltiesAgainst, 
         -faceOffsWonAgainst, -hitsAgainst, -takeawaysAgainst, 
         -giveawaysAgainst, -shotsFor_per_game, -goalsFor_per_game,
         -shotsAgainst_per_game, -goalsAgainst_per_game,
         -penaltiesFor_per_game, -penaltiesAgainst_per_game) %>%
  rename(PTSperc = `P%`, goalsFor_per_game = `GF/GP`,
         goalsAgainst_per_game = `GA/GP`, PP_perc = `PP%`,
         PK_perc = `PK%`, netPP_perc = `Net PP%`, netPK_perc = `Net PK%`,
         shotsFor_per_game = `Shots/GP`, shotsAgainst_per_game = `SA/GP`)
```

```{r}
#adding thru%for and against
just_this_season <- just_this_season %>%
  mutate(shot_thru_perc_for = shotsFor_per_game / shotAttemptsFor_per_game,
         shot_thru_perc_against = 
           shotsAgainst_per_game / shotAttemptsAgainst_per_game)

just_this_season <- just_this_season %>%
  filter(is.na(next_year_pts_perc))
  
```

now to get predictions...

```{r}
tree_predictions <- predict(final_rf_next_year, data = just_this_season,
                             predict.all = TRUE)
```

now I want to randomly sample a prediction from the trees in the forest for each team
```{r}
one_prediction <- just_this_season %>%
  select(team)

pred_locs <- sample(1:500, 31, replace = FALSE)
these_preds <- rep(0, 31)

for(team_i in 1:31) {
  
  this_pred <- tree_predictions$predictions[team_i, pred_locs[team_i]]
  these_preds[team_i] <- this_pred
}

one_prediction <- one_prediction %>%
  bind_cols(predicted_pts_perc = these_preds)

one_prediction <- one_prediction %>%
  mutate(rank = dense_rank(desc(predicted_pts_perc)))

pit_rank <- one_prediction %>%
  filter(team == "PIT") %>%
  select(rank)

```
so that does it once... I want to do it a bunch of times...

```{r}
simulated_pit_ranks <- rep(0, 10000)

set.seed(2020)
for(iter_j in 1:10000) {
  one_prediction <- just_this_season %>%
    select(team)

  pred_locs <- sample(1:500, 31, replace = FALSE)
  these_preds <- rep(0, 31)

  for(team_i in 1:31) {
  
    this_pred <- tree_predictions$predictions[team_i, pred_locs[team_i]]
    these_preds[team_i] <- this_pred
  }

  one_prediction <- one_prediction %>%
    bind_cols(predicted_pts_perc = these_preds)

  one_prediction <- one_prediction %>%
    mutate(rank = rank(desc(predicted_pts_perc), ties.method = "random"))

  pit_rank <- one_prediction %>%
    filter(team == "PIT") %>%
    select(rank)
  
  simulated_pit_ranks[iter_j] <- as.numeric(pit_rank)

}

simulated_pit_ranks <- as_tibble(simulated_pit_ranks) %>%
  rename(rank = value)

```
 so now I have 10000 simulations of the season, determining what rank the Penguins finished in for each one...
 
I kind of want to see what that distribution looks like
```{r}
summary(simulated_pit_ranks)
```

```{r}
simulated_pit_ranks %>%
  ggplot(aes(x = rank)) +
  geom_histogram(bins = 31) +
  theme_bw()
```
now to calculate the probability that the Penguins finish in each spot in the standings...
```{r}
possible_ranks <- c(1:31)
probs <- rep(0, 31)

for(rank_i in 1:31) {
  probs[rank_i] <- sum(simulated_pit_ranks$rank == rank_i) / 10000
}

end_rank_probabilities <- bind_cols(possible_ranks = possible_ranks,
                                    probability = probs)

```

so that creates a dataframe showing the probability of each rank being achieved

we also want to combine these probabilities with the draft lottery probabilities, so to do this we are going to simulate the draft a bunch of times

```{r}

many_predictions <- just_this_season %>%
  select(team)

set.seed(2020)
for(iter_j in 1:10000) {
  one_prediction <- just_this_season %>%
    select(team)

  pred_locs <- sample(1:500, 31, replace = FALSE)
  these_preds <- rep(0, 31)

  for(team_i in 1:31) {
  
    this_pred <- tree_predictions$predictions[team_i, pred_locs[team_i]]
    these_preds[team_i] <- this_pred
  }

  one_prediction <- one_prediction %>%
    bind_cols(predicted_pts_perc = these_preds)

  one_prediction <- one_prediction %>%
    mutate(rank = rank(desc(predicted_pts_perc), ties.method = "random"))
  
  many_predictions <- many_predictions %>%
    bind_cols(predicted_rank = one_prediction$rank)
    

}
```
so that leaves me with a dataset of the 10000 season outcomes, with each team's rankings... now I need to simulate the draft for each of these 10000 season outcomes

```{r}
draft_pick_order_simulations <- just_this_season %>%
  select(team)

pick_order <- rep(0, 31)

simulated_pit_picks <- rep(0, 10000)

set.seed(2021)
for(iter_j in 1:10000) {
  one_simulation <- just_this_season %>%
    select(team)

  this_season_outcome <- many_predictions %>%
    select(iter_j+1)
  
  colnames(this_season_outcome)[1] <- "tmrank"
    
  this_season_outcome <- this_season_outcome %>%
    bind_cols(team = many_predictions$team)
  
  best_16 <- this_season_outcome %>%
    filter(tmrank < 17) %>%
    mutate(pick_order = 32 - tmrank)
  
  this_draft_order <- best_16
  
  worst_15 <- this_season_outcome %>%
    filter(tmrank >= 17) %>%
    mutate(init_pick_order = 32 - tmrank, pick_order = 0)
  
  #find the number 1 spot:
  who_gets_1 <- sample(1:15, size = 1, replace = FALSE, 
                       prob = c(18.5, 13.5, 11.5, 9.5, 8.5, 7.5, 6.5, 6.0,
                                5.0, 3.5, 3.0, 2.5, 2.0, 1.5, 1.0))
  
  probabilities <- as.vector(c(18.5, 13.5, 11.5, 9.5, 8.5, 7.5, 6.5, 6.0,
                                5.0, 3.5, 3.0, 2.5, 2.0, 1.5, 1.0))
  
  for (pick_i in 1:3) {
    who_gets_this_pick <- sample(1:(16-pick_i), size = 1, replace = FALSE, 
                       prob = probabilities)
    
    team_getting_pick <- worst_15 %>%
      filter(init_pick_order == who_gets_this_pick)
    team_getting_pick$pick_order <- pick_i
    
    this_draft_order <- this_draft_order %>%
      bind_rows(team_getting_pick)
    
    worst_15 <- worst_15 %>%
      filter(init_pick_order != who_gets_this_pick)
    
    probabilities <- probabilities[-who_gets_this_pick]
  }
  
  worst_15$pick_order <- rank(worst_15$init_pick_order) + 3
  
  this_draft_order <- this_draft_order %>%
    bind_rows(worst_15)
  
  simulated_pit_picks[iter_j] <- this_draft_order %>%
    filter(team == "PIT") %>%
    select(pick_order)
  
  draft_pick_order_simulations <- draft_pick_order_simulations %>%
    bind_cols(predicted_pick_order = this_draft_order$pick_order)

}

simulated_pit_picks <- as_tibble(simulated_pit_picks) %>%
  rename(pick = value)
```

so now I have 10,000 simulated drafts based on 10,000 simulated seasons
what is the probability of the Penguins ending up in a given draft pick spot?

```{r}


summary(simulated_pit_picks)
```
```{r}
as_tibble(simulated_pit_picks) %>%
  ggplot(aes(x = pick)) +
  geom_histogram(bins = 31) +
  theme_bw()
```
now to calculate the probability that the Penguins get each pick...
```{r}


possible_picks <- c(1:31)
probs_picks <- rep(0, 31)

for(pick_spot_i in 1:31) {
  probs_picks[pick_spot_i] <- sum(simulated_pit_picks$pick == pick_spot_i) / 10000
}

pick_order_probabilities <- bind_cols(possible_pick = possible_picks,
                                    probability = probs_picks)

```

so that creates a dataframe showing the probability of each rank being achieved

I want to export that as a csv so I can use it in another script...
```{r}
write_csv(pick_order_probabilities, "../Data/pickOrderProbabilities.csv")
```



