---
title: "predicting_next_season"
output: html_document
---

now we want to use our random forest to create a distribution of possible season outcomes

I need to basically re-create the random forest model from before, so I need to first manipulate the data like it was in the modeling script

```{r}
library(tidyverse)
library(caret)
library(mgcv)
library(ranger)

complex_team_season_data <-
  read_csv("../Data/predictingPpercNextSeason.csv")
complex_team_season_data <- complex_team_season_data %>%
  mutate(xGperc = xGoalsFor / (xGoalsFor + xGoalsAgainst))
```

Now I want to work on modeling next season's win percentage based on this season's stats...
I think I need to do some data manipulation to turn these into per-game stats and remove some of the named data
```{r}
next_year_model_data <- complex_team_season_data %>%
  select(-team, -season, -season_end, -tm_season) %>%
  mutate(xGFor_per_game = xGoalsFor / GP, 
         shotsFor_per_game = shotsOnGoalFor / GP,
         shotAttemptsFor_per_game = shotAttemptsFor / GP,
         goalsFor_per_game = goalsFor / GP,
         penaltiesFor_per_game = penaltiesFor / GP,
         faceoffsWonFor_per_game = faceOffsWonFor / GP,
         hitsFor_per_game = hitsFor / GP,
         takeawaysFor_per_game = takeawaysFor / GP,
         giveawaysFor_per_game = giveawaysFor / GP,
         xGAgainst_per_game = xGoalsAgainst / GP,
         shotsAgainst_per_game = shotsOnGoalAgainst / GP,
         shotAttemptsAgainst_per_game = shotAttemptsAgainst / GP,
         goalsAgainst_per_game = goalsAgainst / GP,
         penaltiesAgainst_per_game = penaltiesAgainst / GP,
         faceoffsWonAgainst_per_game = faceOffsWonAgainst / GP,
         hitsAgainst_per_game = hitsAgainst / GP,
         takeawaysAgainst_per_game = takeawaysAgainst / GP,
         giveawaysAgainst_per_game = giveawaysAgainst / GP) %>%
  select(-GP, -W, -L, -OT, -P, -RW, -ROW, -`S/O Win`, -xGoalsPercentage,
         -xGoalsFor, -shotsOnGoalFor, -shotAttemptsFor, -goalsFor, 
         -penaltiesFor, -faceOffsWonFor, -hitsFor, -takeawaysFor, 
         -giveawaysFor, -xGoalsAgainst, -shotsOnGoalAgainst, 
         -shotAttemptsAgainst, -goalsAgainst, -penaltiesAgainst, 
         -faceOffsWonAgainst, -hitsAgainst, -takeawaysAgainst, 
         -giveawaysAgainst, -shotsFor_per_game, -goalsFor_per_game,
         -shotsAgainst_per_game, -goalsAgainst_per_game,
         -penaltiesFor_per_game, -penaltiesAgainst_per_game) %>%
  rename(PTSperc = `P%`, goalsFor_per_game = `GF/GP`,
         goalsAgainst_per_game = `GA/GP`, PP_perc = `PP%`,
         PK_perc = `PK%`, netPP_perc = `Net PP%`, netPK_perc = `Net PK%`,
         shotsFor_per_game = `Shots/GP`, shotsAgainst_per_game = `SA/GP`) %>%
  filter(!is.na(next_year_pts_perc))
  
#adding thru%for and against
next_year_model_data <- next_year_model_data %>%
  mutate(shot_thru_perc_for = shotsFor_per_game / shotAttemptsFor_per_game,
         shot_thru_perc_against = 
           shotsAgainst_per_game / shotAttemptsAgainst_per_game)
```



I want to run a number of models, tuning each as needed

first I need to create the folds for 10-fold cross validation
```{r}
set.seed(2020)

train_i <- createDataPartition(y = next_year_model_data$next_year_pts_perc,
                               p = 0.9, list = FALSE) %>%
  as.numeric()

train_nextyear_data <- next_year_model_data[train_i,]
test_nextyear_data <- next_year_model_data[-train_i,]

```

then build the model...
Random Forest
```{r}
rf_tune_grid <-
  expand.grid(mtry = seq(3, 18, by = 3),
              splitrule = "variance",
              min.node.size = 5)

set.seed(2010)
init_rf_nextyear_train <-
  train(next_year_pts_perc ~ ., data = train_nextyear_data,
        method = "ranger", num.trees = 50, 
        trControl = trainControl(method = "cv", number = 5),
        tuneGrid = rf_tune_grid)
plot(init_rf_nextyear_train)
```

```{r}
init_rf_nextyear_train$bestTune
```

```{r}
init_rf_nextyear_train$results[3,]
```


so now I want to build a model using the ranger package and these tuned parameters
```{r}
final_rf_next_year <- ranger(next_year_pts_perc ~ .,
                             data = next_year_model_data,
                             splitrule = "variance",
                             min.node.size = 5,
                             mtry = 9,
                             num.trees = 50)
```

so I have the model, and now I need to get predictions from this model for the 2020-21 season

first I want a dataset that is just the current season...
```{r}
just_this_season <- complex_team_season_data

just_this_season <- just_this_season %>%
  mutate(xGFor_per_game = xGoalsFor / GP, 
         shotsFor_per_game = shotsOnGoalFor / GP,
         shotAttemptsFor_per_game = shotAttemptsFor / GP,
         goalsFor_per_game = goalsFor / GP,
         penaltiesFor_per_game = penaltiesFor / GP,
         faceoffsWonFor_per_game = faceOffsWonFor / GP,
         hitsFor_per_game = hitsFor / GP,
         takeawaysFor_per_game = takeawaysFor / GP,
         giveawaysFor_per_game = giveawaysFor / GP,
         xGAgainst_per_game = xGoalsAgainst / GP,
         shotsAgainst_per_game = shotsOnGoalAgainst / GP,
         shotAttemptsAgainst_per_game = shotAttemptsAgainst / GP,
         goalsAgainst_per_game = goalsAgainst / GP,
         penaltiesAgainst_per_game = penaltiesAgainst / GP,
         faceoffsWonAgainst_per_game = faceOffsWonAgainst / GP,
         hitsAgainst_per_game = hitsAgainst / GP,
         takeawaysAgainst_per_game = takeawaysAgainst / GP,
         giveawaysAgainst_per_game = giveawaysAgainst / GP) %>%
  select(-GP, -W, -L, -OT, -P, -RW, -ROW, -`S/O Win`, -xGoalsPercentage,
         -xGoalsFor, -shotsOnGoalFor, -shotAttemptsFor, -goalsFor, 
         -penaltiesFor, -faceOffsWonFor, -hitsFor, -takeawaysFor, 
         -giveawaysFor, -xGoalsAgainst, -shotsOnGoalAgainst, 
         -shotAttemptsAgainst, -goalsAgainst, -penaltiesAgainst, 
         -faceOffsWonAgainst, -hitsAgainst, -takeawaysAgainst, 
         -giveawaysAgainst, -shotsFor_per_game, -goalsFor_per_game,
         -shotsAgainst_per_game, -goalsAgainst_per_game,
         -penaltiesFor_per_game, -penaltiesAgainst_per_game) %>%
  rename(PTSperc = `P%`, goalsFor_per_game = `GF/GP`,
         goalsAgainst_per_game = `GA/GP`, PP_perc = `PP%`,
         PK_perc = `PK%`, netPP_perc = `Net PP%`, netPK_perc = `Net PK%`,
         shotsFor_per_game = `Shots/GP`, shotsAgainst_per_game = `SA/GP`)
```

```{r}
#adding thru%for and against
just_this_season <- just_this_season %>%
  mutate(shot_thru_perc_for = shotsFor_per_game / shotAttemptsFor_per_game,
         shot_thru_perc_against = 
           shotsAgainst_per_game / shotAttemptsAgainst_per_game)

just_this_season <- just_this_season %>%
  filter(is.na(next_year_pts_perc))
  
```

now to get predictions...

```{r}
tree_predictions <- predict(final_rf_next_year, data = just_this_season,
                             predict.all = TRUE)
```

now I want to randomly sample a prediction from the trees in the forest for each team
```{r}
one_prediction <- just_this_season %>%
  select(team)

pred_locs <- sample(1:50, 31, replace = FALSE)
these_preds <- rep(0, 31)

for(team_i in 1:31) {
  
  this_pred <- tree_predictions$predictions[team_i, pred_locs[team_i]]
  these_preds[team_i] <- this_pred
}

one_prediction <- one_prediction %>%
  bind_cols(predicted_pts_perc = these_preds)

one_prediction <- one_prediction %>%
  mutate(rank = dense_rank(desc(predicted_pts_perc)))

pit_rank <- one_prediction %>%
  filter(team == "PIT") %>%
  select(rank)

```
so that does it once... I want to do it a bunch of times...

```{r}
simulated_pit_ranks <- rep(0, 10000)

for(iter_j in 1:10000) {
  one_prediction <- just_this_season %>%
    select(team)

  pred_locs <- sample(1:50, 31, replace = FALSE)
  these_preds <- rep(0, 31)

  for(team_i in 1:31) {
  
    this_pred <- tree_predictions$predictions[team_i, pred_locs[team_i]]
    these_preds[team_i] <- this_pred
  }

  one_prediction <- one_prediction %>%
    bind_cols(predicted_pts_perc = these_preds)

  one_prediction <- one_prediction %>%
    mutate(rank = dense_rank(desc(predicted_pts_perc)))

  pit_rank <- one_prediction %>%
    filter(team == "PIT") %>%
    select(rank)
  
  simulated_pit_ranks[iter_j] <- as.numeric(pit_rank)

}

simulated_pit_ranks <- as_tibble(simulated_pit_ranks) %>%
  rename(rank = value)

```
 so now I have 10000 simulations of the season, determining what rank the Penguins finished in for each one...
 
I kind of want to see what that distribution looks like
```{r}
summary(simulated_pit_ranks)
```

```{r}
simulated_pit_ranks %>%
  ggplot(aes(x = rank)) +
  geom_histogram(bins = 31) +
  theme_bw()
```
now to calculate the probability that the Penguins finish in each spot in the standings...
```{r}
possible_ranks <- c(1:31)
probs <- rep(0, 31)

for(rank_i in 1:31) {
  probs[rank_i] <- sum(simulated_pit_ranks$rank == rank_i) / 10000
}

end_rank_probabilities <- bind_cols(possible_ranks = possible_ranks,
                                    probability = probs)

```

so that creates a dataframe showing the probability of each rank being achieved